### 2015-08-23 output/table_meancorr_laborm_patentm.tex

\begin{table}
\begin{small}
\begin{threeparttable}
\caption{{\normalsize Average correlations of labor market outcomes and automation across industries}}
\label{tab:meancorr_laborm_patentm}
\begin{tabular}{lrrrrrrrrrr}
\toprule \addlinespace[0.5em]
 & \rot{Employment} & \rot{Employment share} &  \rot{Capital productivity} & \rot{Output} & \rot{Capital cost} &  \rot{Labor cost} & \rot{Output deflator} & \rot{Production} & \rot{Capital} & \rot{Labor productivity} \tabularnewline[0.05cm]
\midrule \addlinespace[0.5em]
\# Automation patents$^1$ & \textbf{-0.56} & -0.02 & 0.04 & 0.36 & 0.35 & 0.46 & \textbf{0.49} & \textbf{0.52} & 0.6 & \textbf{0.76}  \tabularnewline[0.0cm]
 & \scriptsize{\scriptsize{(0.46)}} & \scriptsize{\scriptsize{(0.46)}} & \scriptsize{(0.41)} & \scriptsize{(0.59)} & \scriptsize{(0.44)} & \scriptsize{(0.50)} & \scriptsize{(0.23)} & \scriptsize{(0.38)} & \scriptsize{(0.62)} & \scriptsize{(0.65)} \tabularnewline[0.1cm]
Average match per patent & \textbf{-0.59} & 0 & 0.1 & 0.33 & 0.4 & 0.46 & \textbf{0.52} & \textbf{0.55} & 0.56 & \textbf{0.77}  \tabularnewline[0.0cm]
 & \scriptsize{\scriptsize{(0.48)}} & \scriptsize{\scriptsize{(0.50)}} & \scriptsize{(0.38)} & \scriptsize{(0.51)} & \scriptsize{(0.44)} & \scriptsize{(0.52)} & \scriptsize{(0.21)} & \scriptsize{(0.32)} & \scriptsize{(0.63)} & \scriptsize{(0.69)} \tabularnewline[0.1cm]
Share of automation patents & \textbf{-0.67} & -0.02 & 0.14 & 0.34 & 0.4 & 0.48 & \textbf{0.59} & \textbf{0.6} & 0.55 & \textbf{0.87}  \tabularnewline[0.0cm]
 & \scriptsize{\scriptsize{(0.49)}} & \scriptsize{\scriptsize{(0.52)}} & \scriptsize{(0.44)} & \scriptsize{(0.53)} & \scriptsize{(0.45)} & \scriptsize{(0.53)} & \scriptsize{(0.17)} & \scriptsize{(0.37)} & \scriptsize{(0.66)} & \scriptsize{(0.74)} \tabularnewline[0.1cm]
Automation index$^2$ & \textbf{-0.59} & 0 & 0.1 & 0.33 & 0.4 & 0.46 & \textbf{0.52} & \textbf{0.55} & 0.56 & \textbf{0.77}  \tabularnewline[0.0cm]
 & \scriptsize{\scriptsize{(0.48)}} & \scriptsize{\scriptsize{(0.50)}} & \scriptsize{(0.38)} & \scriptsize{(0.51)} & \scriptsize{(0.44)} & \scriptsize{(0.52)} & \scriptsize{(0.21)} & \scriptsize{(0.32)} & \scriptsize{(0.63)} & \scriptsize{(0.69)} \tabularnewline[0.1cm]
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Note:} Correlations are calculated from the yearly time series 1987-2014 of the automation indicator and the labor market outcome series. The values are cross-sectional averages across the 26 manufacturing industries. Values in parentheses show standard deviations. Bold values are at least one standard deviation away from 0.
\item $^1$: Automation patents are defined as patents with at least one keyword match.
\item $^2$: Calculated as $log(1 + \text{number of matches per patent})$.
\item \textit{Source:} BLS, USPTO, Google and own calculations.
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{table}